Introduction to Markov state modeling with the PyEMMA software --- v\versionnumber

### 2018-11-14 manuscript/manuscript.tex

In Markov state modeling, our objective is to model the slow dynamics of a molecular process.
In order to approximate the slow dynamics in a statistically efficient manner,
a lower dimensional representation of our simulation data is necessary.
However, the features (e.g. torsion angles, distances or contacts) which best represent the slow dynamical modes of a given molecular system are unknown a priori~\cite{NoeClementiReview}.
Fortunately, the variational principle of conformational dynamics~\cite{noe-vac,nueske-vamk}
and the more general variational approach for Markov processes (VAMP)~\cite{vamp-preprint}
provide a systematic means to quantitatively compare multiple representations of the simulation data.
In particular, we can use a scalar score obtained using VAMP to directly compare the ability of certain features to capture slow dynamical modes in a particular molecular system.

Here, we utilize the VAMP-2 score, which maximizes the kinetic variance contained in the features~\cite{kinetic-maps}.
We should always evaluate the score in a cross-validated manner to ensure that we neither include too few features (under-fitting) or too many features (over-fitting)~\cite{gmrq,vamp-preprint}.
To choose among three different molecular features reflecting protein structure,
we compute the (cross-validated) VAMP-2 score (notebook 00).
Although we cannot MSM optimize lag times with a variational score\cite{husic2017note}, such as VAMP-2,
it is important to ensure that properties that we optimize are robust as a function of lag time. 
Consequently, we compute the VAMP-2 score at several lag times (notebook 00). 
We find that the relative rankings of the different molecular features are highly robust as a function of lag time. 
We show one example of this ranking and the absolute VAMP-2 scores for lag time~$0.5$~ns in Fig.~\ref{fig:io-to-tica}b. 
We find that backbone torsions contain more kinetic variance than the backbone heavy atom positions or the distances between them (Fig.~\ref{fig:io-to-tica}b).
This suggests that backbone torsions are the best of the options evaluated for MSM construction.

@@

@@

@@

More details on how to model complex systems with the techniques presented here are described, e.g., by~\cite{plattner_protein_2015,plattner_complete_2017}.

>>>>

In Markov state modeling, our objective is to model the slow dynamics of a molecular process.
In order to approximate the slow dynamics in a statistically efficient manner,
a lower dimensional representation of our simulation data is necessary.
However, the features (e.g. torsion angles, distances or contacts) which best represent the slow dynamical modes of a given molecular system are unknown a priori~\cite{NoeClementiReview}.
Fortunately, the variational principle of conformational dynamics~\cite{noe-vac,nueske-vamk}
and the more general variational approach for Markov processes (VAMP)~\cite{vamp-preprint}
provide a systematic means to quantitatively compare multiple representations of the simulation data.
In particular, we can use a scalar score obtained using VAMP to directly compare the ability of certain features to capture slow dynamical modes in a particular molecular system.
In Notebook (01), we present in detail how to extract features from MD datasets and how to systematically compare them.

Throughout this tutorial, we utilize the VAMP-2 score, which maximizes the kinetic variance contained in the features~\cite{kinetic-maps}.
We should always evaluate the score in a cross-validated manner to ensure that we neither include too few features (under-fitting) or too many features (over-fitting)~\cite{gmrq,vamp-preprint}.
To choose among three different molecular features reflecting protein structure,
we compute the (cross-validated) VAMP-2 score (notebook 00).
Although we cannot MSM optimize lag times with a variational score\cite{husic2017note}, such as VAMP-2,
it is important to ensure that properties that we optimize are robust as a function of lag time. 
Consequently, we compute the VAMP-2 score at several lag times (notebook 00). 
We find that the relative rankings of the different molecular features are highly robust as a function of lag time. 
We show one example of this ranking and the absolute VAMP-2 scores for lag time~$0.5$~ns in Fig.~\ref{fig:io-to-tica}b. 
We find that backbone torsions contain more kinetic variance than the backbone heavy atom positions or the distances between them (Fig.~\ref{fig:io-to-tica}b).
This suggests that backbone torsions are the best of the options evaluated for MSM construction.

@@

We demonstrate how to apply TICA, suggest how to interpret the projected coordinates, and compare the results to other dimension reduction techniques in Notebook (02).

@@

In addition to a detailed demonstration of the above, Notebook (06) demonstrates how to compute J-couplings and dynamic fingerprints from MSMs.

@@

More details on how to model complex systems with the techniques presented here are described, e.g., by~\cite{plattner_protein_2015,plattner_complete_2017}.
We further examine some symptoms that may indicate problematic or difficult datasets, and demonstrate how to deal with them in Notebook (08).
